---
title: "Term Counts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
source('mesh_helpers.R')
```


```{r}
term_counts = loadMeshTermsCounts('data/mesh_term_counts.RDS')
head(term_counts)
```


```{r}
mesh_levels = read.csv('data/desc2020_compressed.csv')
head(mesh_levels)
```


##View Count Summaries for Level and Type

```{r}
term_counts %>% nrow %>% print
mesh_levels %>% nrow %>% print
```

```{r}
term_counts %>% 
  filter(term %in% mesh_levels$Descriptor) %>%
  group_by(type) %>%
  summarise(n = n())
```

No qualifiers are found, but this may be because the mesh levels are Title Case and the qualifiers are lower case.
```{r}
term_counts$TermUpper = toupper(term_counts$term)
mesh_levels$DescriptorUpper = toupper(mesh_levels$Descriptor)

term_counts %>%
  filter(TermUpper %in% mesh_levels$DescriptorUpper) %>%
  group_by(type) %>%
  summarise(n = n(), total = sum(term_counts$type == type), percent = as.integer(n/total * 100))
```

```{r}
term_counts %>%
  filter(TermUpper %in% mesh_levels$DescriptorUpper) %>%
  summarise(n = n(), total = nrow(term_counts), percent = as.integer(n/total * 100))
```

* Type Descriptor (99%) and type Qualifier (44%) are in the mesh levels
* We have IDs for 21455 of the 21563 terms in term counts 

Examine the levels that are not found in our database of mesh terms
```{r}
mesh_levels$Found = mesh_levels %>% 
  select(DescriptorUpper) %>% 
  sapply(function(t) t %in% term_counts$TermUpper)
mesh_levels %>% filter(!Found) %>% sample_n(10)

mesh_levels$Level = sapply(mesh_levels$TreeNumber, function(tn) substr(tn, 1, 1))
```

Examine found vs not found by level
```{r}
mesh_levels %>% group_by(Level, Found) %>% summarise(n = n()) %>%
  ggplot(aes(Level, n)) + 
  geom_bar(aes(fill = Found), stat="identity", position = "dodge")
```


##Approach by Term Counts 

Compare group-wise frequency differences between terms in mesh levels and terms outside mesh levels.
```{r}
term_counts$Found = term_counts$TermUpper %>% 
  sapply(function(t) t %in% mesh_levels$DescriptorUpper)

term_counts$LogN = log10(term_counts$n)
```

```{r}
ggplot(term_counts, aes(Found, LogN)) + geom_boxplot()
```

```{r}
term_counts %>% 
  group_by(Found) %>%
  summarise(minN = min(n), maxN = max(n), maxLogN = max(LogN))
```

```{r}
term_counts %>% 
  filter(Found) %>% 
  {ggplot(.) + geom_density(aes(x = LogN))}
```


```{r}
term_counts %>%
  filter(!Found) %>%
  {ggplot(.) + geom_density(aes(x = LogN))}
```

```{r}
term_counts_found = merge(term_counts, mesh_levels, by.x = "TermUpper", by.y = "DescriptorUpper")
nrow(term_counts_found)
```

A lot of terms were found more than once.


```{r}
term_counts_found_unique = term_counts_found %>% 
  group_by(TermUpper) %>%
  summarise(TreeNumber = first(TreeNumber), Level = first(Level),
            term = first(term), UI = first(UI),
            `2011` = first(`2011`), `2012` = first(`2012`),
            `2013` = first(`2013`), `2014` = first(`2014`),
            `2015` = first(`2015`), `2016` = first(`2016`),
            `2017` = first(`2017`), n = first(n),
            percent_with_term = first(percent_with_term)) %>% 
  as.data.frame

head(term_counts_found_unique)
```

```{r}
nrow(term_counts_found_unique)
```

```{r}
term_counts_found_unique %>% 
  group_by(Level) %>% summarise(n = n()) %>%
  ggplot(aes(Level, n)) + 
  geom_bar(stat="identity", position = "dodge")
```

Write to file
```{r}
saveRDS(term_counts_found_unique, 'data/mesh_term_counts_ided.RDS')
```

```{r}
term_counts_found_unique %>% 
  filter(substr(TreeNumber, 1, 1) == "C") %>% 
  saveRDS('data/mesh_term_counts_ided_c_only.RDS')
```